---
title: "p8105_hw6_lz2950"
author: "Lehan Zou"
date: "2023-12-03"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(ggplot2)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Due date

Due: December 3 at 11:59pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |


### Problem 0

This "problem" focuses on structure of your assignment, including the use of R Markdown to write reproducible reports, the use of R Projects to organize your work, the use of relative paths to load data, and the naming structure for your files. 

To that end: 

* create a public GitHub repo + local R Project for this assignment
* write solutions using a .Rmd file that outputs a `github_document` / .md file
* submit a link to your repo via Courseworks

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems. 

For Problem 0, we will assess adherence to the instructions above regarding repo structure, git commit history, and whether we are able to knit your .Rmd to ensure that your work is reproducible. Adherence to appropriate styling and clarity of code will be assessed in Problems 1+ using the homework [style rubric](homework_style_rubric.html). 

This homework includes figures; the readability of your embedded plots (e.g. font sizes, axis labels, titles) will be assessed in Problems 1+.

```{r load_libraries}
library(tidyverse)
library(modelr)
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2

**The boostrap is helpful when you'd like to perform inference for a parameter / value / summary that doesn't have an easy-to-write-down distribution in the usual repeated sampling framework. We'll focus on a simple linear regression with `tmax` as the response with `tmin` and `prcp` as the predictors, and are interested in the distribution of two quantities estimated from these data:**

-   $\hat{r}^2$

-   $log(\hat{\beta_1} * \hat{\beta_2})$

**Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution.

First, download the data.
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}

boot_strap = weather_df |>
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    r_square = map(models, broom::tidy),
    beta = map(models, broom::glance))|> 
  unnest(r_square, beta) |>
  select(.id, term, estimate, r.squared) |>
  pivot_wider(names_from = term, values_from = estimate)

var = boot_strap |>
  summarise(r.squared = r.squared,
    log_beta12 = log(tmin * prcp))
```


```{r}
var |> 
  ggplot(aes(x = r.squared)) +
  geom_density() +
  labs(title = "r squared estimates Distribution",
       x = "r square")
```
This plot of r square is left skewed with the mode around 0.92.But this distribution is approximately normal. 
```{r}
var |>
  ggplot(aes(x = log_beta12)) +
  geom_density() +
  labs(title = "log_beta1 * log_beta2 Estimate Distribution",
       x = "log_beta1 * log_beta2")

```
This plot of log_beta1*log_beta2 is also left skewed with some outliers.

```{r}
tibble(
  r_square_low = var |> pull(r.squared) |> quantile(0.025),
  r_square_high = var |> pull(r.squared) |> quantile(0.975),
  log_beta_low = var |> pull(log_beta12) |> quantile(0.025, na.rm = T),
  log_beta_high = var |> pull(log_beta12) |> quantile(0.975, na.rm = T)
) |> 
  knitr::kable()
```

### Problem 3
```{r}
birthwt =
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate( 
    babysex = factor(
      case_match(
        babysex,
        1 ~ "male",
        2 ~ "female"
      )
    ),
    frace = factor(
      case_match(
        frace,
        1 ~ "White",
        2 ~ "Black",
        3 ~ "Asian",
        4 ~ "Puerto Rican",
        8 ~ "Other",
        9 ~ "Unknown"
      )
    ),
    mrace = factor(
      case_match(
        mrace,
        1 ~ "White",
        2 ~ "Black",
        3 ~ "Asian",
        4 ~ "Puerto Rican",
        8 ~ "Other"
      )
    ),
    malform = factor(
      case_match(
        malform,
        0 ~ "absent",
        1 ~ "present"
      )
    )
  )

# check NAs
sum(is.na(birthwt))
```
The resulting data set has `r nrow(birthwt)` observations and `r ncol(birthwt)` variables. There is no NA values in this data set.

Then I fitted the data to a model with all the variables included as predictors because I want to see the overall performance first.
```{r}
birthwt_model =
  birthwt |> 
  lm(bwt ~ ., data = _)

birthwt |> 
  modelr::add_predictions(birthwt_model) |> 
  modelr::add_residuals(birthwt_model) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point()
```
From the plot, it seems no strong relation between residuals and fitted values.

Compare the model to two others:

 * First: One using length at birth and gestational age as predictors (main effects only)

 * Second: One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
cv_df = 
  crossv_mc(birthwt, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) |> 
  mutate(
    model  = map(train, \(df) lm(bwt ~ ., data = df)),
    model1  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model2  = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead * blength + blength * babysex + bhead * babysex + bhead * blength * babysex, data = df))) |> 
  mutate(
    rmse = map2_dbl(model, test, \(mod, df) rmse(model = mod, data = df)),
    rmse1 = map2_dbl(model1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse2 = map2_dbl(model2, test, \(mod, df) rmse(model = mod, data = df))
  )

```

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

We can conclude from the plot that the original model with all the variables as predictors has the smallest mse. This concludes that the number of variables and interactions between variables is worth to take into consideration. 









